Loss distributions
==================

```{r, include=FALSE}
tutorial::go_interactive()
```

## Introduction


The aim of the course is to provide a fundamental basis which applies
mainly in general insurance. General insurance companies' products are
short-term policies that can be purchased for a short period of time.
Examples of insurance products are

-   motor insurance;

-   home insurance;

-   health insurance; and

-   travel insurance.

In case of an occurrence of an insured event, two important components
of financial losses which are of importance for management of an
insurance company are

-   the number of claims; and

-   the amounts of those claims.

Mathematical and statistical techniques used to model these sources of
uncertainty will be discussed. This will enable insurance companies to

-   calculate premium rates to charge policy holders; and

-   decide how much reserve should be set aside for the future payment
    of incurred claims.

In the chapter, statistical distributions and their properties which are
suitable for modelling claim sizes are reviewed. These distribution are
also known as loss distributions. In practice, the shape of loss
distributions are positive skew with a long right tail. The main
features of loss distributions include:

-   having a few small claims;

-   rising to a peak;

-   tailing off gradually with a few very large claims.

Exponential Distribution
------------------------

A random variable $X$ has an exponential distribution with a parameter
$\lambda > 0$, denoted by $X \sim \text{Exp}(\lambda)$ if its
probability density function is given by
$$f_X(x) = \lambda e^{-\lambda x}, \quad x > 0.$$

::: {.example}
*Let $X \sim \text{Exp}(\lambda)$ and $0 < a < b$.*

1.  *Find the distribution $F_X(x)$.*

2.  *Express $P(a < X < B)$ in terms of $f_X(x)$ and $F_X(x)$.*

3.  *Show that the moment generating function of $X$ is
    $$M_X(t) = \left(1 -  \frac{t}{\lambda}\right)^{-1}, \quad t < \lambda.$$*

4.  *Derive the $r$-th moment about the origin $\mathrm{E}[X^r].$*

5.  *Derive the coefficient of skewness for $X$.*

6.  *Simulate a random sample of size n = 200 from
    $X \sim \text{Exp}(0.5)$ using the command
    `sample = rexp(n, rate = lambda)` where $n$ and $\lambda$ are the chosen
    parameter values.*

7.  *Plot a histogram of the random sample using the command
    `hist(sample)` (use help for available options for `hist` function
    in R).*
:::
**Solution:**
The code for questions 6 and 7 is given below. The histogram can be generated from the code below.

``` {frame="single" caption="R code for Example1(6-7)."}
# set.seed is used so that random number generated from different simulations are the same. 
# The number 5353 can be set arbitrarily. 
set.seed(5353)

nsample <- 200
data_exp <- rexp(nsample, rate = 0.5)

dataset <- data_exp
hist(dataset, breaks=100,probability = TRUE, xlab = "claim sizes" 
     , ylab = "density", main = paste("Histogram of claim sizes" ))

hist(dataset, breaks=100, xlab = "claim sizes" 
     , ylab = "count", main = paste("Histogram of claim sizes" ))
```

<!-- ![ **The histogram of the simulated random samples from -->
<!-- $X \sim \text{Exp}(0.5)$** ](FigHistogramExp.eps){#FigHistogramExp -->
<!-- width="80%"} -->



<script src=https://cdn.datacamp.com/datacamp-light-latest.min.js></script>

Copy and paste the code above and run it.

```{r}
set.seed(5353)

nsample <- 200
data_exp <- rexp(nsample, rate = 0.5)

dataset <- data_exp
hist(dataset, breaks=100,probability = TRUE, xlab = "claim sizes" 
     , ylab = "density", main = paste("Histogram of claim sizes" ))


```


```{r}
set.seed(5353)

nsample <- 200
data_exp <- rexp(nsample, rate = 0.5)

dataset <- data_exp


hist(dataset, breaks=100, xlab = "claim sizes" 
     , ylab = "count", main = paste("Histogram of claim sizes" ))

```



**Notes**

1.  The exponential distribution can used to model the inter-arrival
    time of an event.

2.  The exponential distribution has an important property called **lack
    of memory**: if $X \sim \text{Exp}(\lambda)$, then the random
    variable $X-w$ conditional on $X > w$ has the same distribution as
    $X$, i.e.
    $$X \sim \text{Exp}(\lambda)\Rightarrow  X - w | X > w \sim \text{Exp}(\lambda).$$


We can use R to plot the probability density functions (pdf) of exponential distributions with various parameters $\lambda$, which are shown in  Figure \@ref(fig:FigExp). Here we use `scale_colour_manual` to override defaults with scales package (see cheat sheet for details).


```{r FigExp, tut = FALSE, fig.cap= 'The probability density functions (pdf) of exponential distributions with various parameters lambda.'}
library(ggplot2)
ggplot(data.frame(x=c(0,10)), aes(x=x)) + 
  labs(y="Probability density", x = "x") +
  ggtitle("Exponential distributions") +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_function(fun=dexp,geom ="line", args = (mean=0.5), aes(colour = "0.5")) +
  stat_function(fun=dexp,geom ="line", args = (mean=1), aes(colour = "1")) +
  stat_function(fun=dexp,geom ="line", args = (mean=1.5), aes(colour = "1.5")) +
  stat_function(fun=dexp,geom ="line", args = (mean=2), aes(colour = "2")) + 
  scale_colour_manual(expression(paste(lambda, " = ")), values = c("red", "blue", "green", "orange"))
```




Gamma distribution
------------------

A random variable $X$ has a gamma distribution with parameters
$\alpha > 0$ and $\lambda > 0$, denoted by
$X \sim \mathcal{G}(\alpha, \lambda)$ or
$X \sim \text{gamma}(\alpha, \lambda)$ if its probability density
function is given by
$$f_X(x) = \frac{\lambda^\alpha}{\Gamma(\alpha)} x^{\alpha -1} e^{-\lambda x}, \quad x > 0.$$
The symbol $\Gamma$ denotes the gamma function, which is defined as
$$\Gamma(\alpha) = \int_{0}^\infty x^{\alpha - 1} e^{-x} \mathop{}\!dx, \quad \text{for } \alpha > 0.$$
It follows that $\Gamma(\alpha + 1) = \alpha \Gamma(\alpha)$ and that
for a positive integer $n$, $\Gamma(n) = (n-1)!$.

The properties of the gamma distribution are summarised.

-   The mean and variance of $X$ are
    $$\mathrm{E}[X] = \frac{\alpha}{\lambda} \text{ and } \mathrm{Var}[X] =\frac{\alpha}{\lambda^2}$$

-   The $r$-th moment about the origin is
    $$\mathrm{E}[X^r] = \frac{1}{\lambda^r} \frac{\Gamma(\alpha + r)}{\Gamma(\alpha )}, \quad r > 0.$$

-   The moment generating function (mgf) of $X$ is
    $$M_X(t) = \left(1 -  \frac{t}{\lambda}\right)^{-\alpha}, \quad t < \lambda.$$

-   The coefficient of skewness is $$\frac{2}{\sqrt{\alpha}}.$$

**Notes**
1.  The exponential function is a special case of the gamma
    distribution, i.e. $\text{Exp}(\lambda)= \mathcal{G}(1,\lambda)$

2.  If $\alpha$ is a positive integer, the sum of $\alpha$ independent,
    identically distributed as $\text{Exp}(\lambda)$, is
    $\mathcal{G}(\alpha, \lambda)$.

3.  If $X_1, X_2, \ldots, X_n$ are independent, identically distributed,
    each with a $\mathcal{G}(\alpha, \lambda)$ distribution, then
    $$\sum_{i = 1}^n X_i \sim \mathcal{G}(n\alpha, \lambda).$$

4.  The exponential and gamma distributions are not fat-tailed, and
    **may not provide a good fit** to claim amounts.

::: {.example}
*Using the moment generating function of a gamma
distribution, show that the sum of independent gamma random variables
with the same scale parameter $\lambda$,
$X \sim \mathcal{G}(\alpha_1, \lambda)$ and
$Y \sim \mathcal{G}(\alpha_2, \lambda)$, is
$S = X+ Y \sim \mathcal{G}(\alpha_1 + \alpha_2, \lambda).$*
:::

**Solution:**
Because $X$ and $Y$ are independent, $$\begin{aligned}
    M_S(t) &= M_{X+Y}(t) = M_X(t) \cdot M_Y(t)\\
        &= (1 - \frac{t}{\lambda})^{-\alpha_1} \cdot (1 - \frac{t}{\lambda})^{-\alpha_2}     \\
        &=  (1 - \frac{t}{\lambda})^{-(\alpha_1 + \alpha_2)}. \end{aligned}$$
Hence $S = X + Y \sim \mathcal{G}(\alpha_1 + \alpha_2, \lambda).$


<!-- ![ **The probability density functions (pdf) of gamma distributions with -->
<!-- various shape parameters $\alpha$ and rate parameter $\lambda = 1$.** -->
<!-- ](FigGamma.eps){width="80%"} -->

<!-- ![ **The probability density functions (pdf) of lognormal distributions -->
<!-- with $\mu = 0$ and $\sigma = 0.25$ or 1.** -->
<!-- ](FigLognormal.eps){width="80%"} -->


The probability density functions (pdf) of gamma distributions with
various shape parameters $\alpha$ and rate parameter $\lambda$ = 1 are shown in Figure \@ref(fig:FigGamma).



```{r FigGamma, tut = FALSE, fig.cap= 'The probability density functions (pdf) of gamma distributions with various shape alpha and rate parameter lambda = 1.'}
ggplot(data.frame(x=c(0,20)), aes(x=x)) +
  labs(y="Probability density", x = "x") + 
  ggtitle("Gamma distribution")  +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_function(fun=dgamma, args=list(shape=2, rate=1), aes(colour = "2")) +
  stat_function(fun=dgamma, args=list(shape=6, rate=1) , aes(colour = "6")) +
  scale_colour_manual(expression(paste(lambda, " = 1 and ", alpha ," = ")), values = c("red", "blue"))
```




Lognormal distribution
----------------------

A random variable $X$ has a lognormal distribution with parameters $\mu$
and $\sigma^2$, denoted by $X \sim \mathcal{LN}(\mu, \sigma^2)$ if its
probability density function is given by
$$f_X(x) = \frac{1}{\sigma x \sqrt{2 \pi}} \exp\left(-\frac{1}{2} \left( \frac{\log(x) - \mu}{\sigma} \right)^2 \right) , \quad x > 0.$$

The following relation holds:
$$X \sim \mathcal{LN}(\mu, \sigma^2)\Leftrightarrow Y = \log X \sim \mathcal{N}(\mu, \sigma^2).$$

The properties of the lognormal distribution are summarised.

-   The mean and variance of $X$ are
    $$\mathrm{E}[X] = \exp\left(\mu + \frac{1}{2} \sigma^2 \right) \text{ and } \mathrm{Var}[X] =\exp\left(2\mu +  \sigma^2 \right) (\exp(\sigma^2) - 1).$$

-   The $r$-th moment about the origin is
    $$\mathrm{E}[X^r] =\exp\left(r\mu +  \frac{1}{2}r^2 \sigma^2 \right).$$

-   The moment generating function (mgf) of $X$ is not finite for any
    positive value of $t$.

-   The coefficient of skewness is
    $$(\exp(\sigma^2)  + 2) \left(\exp(\sigma^2)  -1 \right)^{1/2} .$$

The probability density functions (pdf) of gamma distributions with
various shape parameters $\alpha$ and rate parameter $\lambda = 1$ is shown in Figure \@ref(fig:FigLognormal).

```{r FigLognormal, tut = FALSE, fig.cap= 'The probability density functions (pdf) of lognormal distributions with mu = 0 and sigma = 0.25 or 1.'}
ggplot(data.frame(x=c(0,10)), aes(x=x)) +
  labs(y="Probability density", x = "x") + 
  ggtitle("lognormal distribution")  +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_function(fun=dlnorm, args = list(meanlog = 0, sdlog = 0.25), aes(colour = "0.25")) +
  stat_function(fun=dlnorm, args = list(meanlog = 0, sdlog = 1), aes(colour = "1")) +
  scale_colour_manual(expression(paste(mu, " = 0 and ", sigma, "= ")), values = c("red", "blue"))

```


Pareto distribution
-------------------

A random variable $X$ has a Pareto distribution with parameters
$\alpha > 0$ and $\lambda > 0$, denoted by
$X \sim \text{Pa}(\alpha, \lambda)$ if its probability density function
is given by
$$f_X(x) = \frac{\alpha \lambda^\alpha}{(\lambda + x)^{\alpha + 1}}, \quad x > 0.$$
The distribution function is given by
$$F_X(x) = 1 - \left(  \frac{\lambda}{\lambda + \alpha} \right)^\alpha, \quad x > 0.$$

The properties of the Pareto distribution are summarized.

-   The mean and variance of $X$ are
    $$\mathrm{E}[X] = \frac{\lambda}{\alpha - 1}, \alpha > 1 \text{ and } \mathrm{Var}[X] = \frac{\alpha \lambda^2}{(\alpha - 1)^2(\alpha - 2)}, \alpha > 2.$$

-   The $r$-th moment about the origin is
    $$\mathrm{E}[X^r] =\frac{\Gamma(\alpha-r) \Gamma(1+ r)}{\Gamma(\alpha)} \lambda^r, \quad 0 < r < \alpha.$$

-   The moment generating function (mgf) of $X$ is not finite for any
    positive value of $t$.

-   The coefficient of skewness is
    $$\frac{2(\alpha + 1)}{\alpha - 3} \sqrt{\frac{\alpha-2}{\alpha}} , \quad \alpha > 3.$$

**Note**
1.  The following conditional tail property for a Pareto distribution is
    useful for reinsurance calculation. Let
    $X \sim \text{Pa}(\alpha, \lambda)$. Then the random variable
    $X - w$ conditional on $X > w$ has a Pareto distribution with
    parameters $\alpha$ and $\lambda + w$, i.e.
    $$X \sim \text{Pa}(\alpha, \lambda)\Rightarrow  X - w | X > w \sim \text{Pa}(\alpha,\lambda + w).$$

2.  The lognormal and Pareto distributions, in practice, provide a
    better fit to claim amounts than exponential and gamma
    distributions.

3.  Other loss distribution are useful in practice including **Burr,
    Weibull and loggamma distributions**.




```{r FigPareto, tut = FALSE, fig.cap= 'The probability density functions (pdf) of Pareto distributions with various shape alpha and rate parameter lambda = 1.', message= FALSE}
library(actuar)
ggplot(data.frame(x=c(0,60)), aes(x=x)) +
  labs(y="Probability density", x = "x") + 
  ggtitle("Pareto distribution")  +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_function(fun=dpareto, args=list(shape=3, scale=20), aes(colour = "alpha = 3, lambda = 20")) +
  stat_function(fun=dpareto, args=list(shape=6, scale=50), aes(colour = "alpha = 6, lambda = 50")) +
  scale_colour_manual("Parameters", values = c("red", "blue"), labels = c(expression(paste(alpha, " = 3 and ", lambda, "= 20")),  expression(paste(alpha, " = 6 and ", lambda, "= 50")))) 
```


::: {.example #exampleFittingClaimSizes}
*Consider a data set consisting of 200 claim amounts in
one year from a general insurance portfolio.*

1.  *Calculate the sample mean and sample standard deviation.*

2.  *Use the method of moments to fit these data with both exponential
    and gamma distributions.*

3.  *Calculate the boundaries for groups or bins so that the expected
    number of claims in each bin is 20 under the fitted exponential
    distribution.*

4.  *Count the values of the observed claim amounts in each bin.*

5.  *With these bin boundaries, find the expected number of claims when
    the data are fitted with the gamma,  lognormal and Pareto distributions.*

6.  *Plot a histogram for the data set along with fitted exponential
    distribution and fitted gamma distribution. In addition, plot another histogram for the data set along with fitted lognormal
    and fitted Pareto distribution.*
    
    

7.  *Comment on the goodness of fit of the fitted distributions.*
:::


**Solution:**
1.  Given that $\sum_{i=1}^n x_i = 206046.4$ and
    $\sum_{i=1}^n x_i^2 = 1,472,400,135$, we have
    $$\bar{x} = \frac{\sum_{i=1}^n x_i}{n} = \frac{206046.4}{200} = 1030.232.$$
    The sample variance and standard deviation are
    $$s^2 = \frac{1}{n-1} \left( \sum_{i=1}^n x_i^2 - \frac{(\sum_{i=1}^n x_i)^2}{n} \right) = 6332284,$$
    and $$s = 2516.403.$$

<!-- 2.  [\[example2\]]{#example2 label="example2"} -->

2. We calculate estimates of
    unknown parameters of both exponential and gamma distributions by
    the method of moments. We simply match the mean and central moments,
    i.e. matching $\mathrm{E}[X]$ to the sample mean $\bar{x}$ and
    $\mathrm{Var}[X]$ to the sample variance.

    The MME (moment matching estimation) of the required distributions
    are as follows:

    -   the MME of $\lambda$ for an $\text{Exp}(\lambda)$ distribution
        is the reciprocal of the sample mean,
        $$\tilde{\lambda} = \frac{1}{\bar{x}} = 0.000971.$$

    -   the MMEs of $\alpha$ and $\lambda$ for a
        $\mathcal{G}(\alpha, \lambda)$ distribution are
        $$\begin{aligned}
        \tilde{\alpha} &= \left(\frac{\bar{x}}{s}\right)^2 = 0.167614, \\
        \tilde{\lambda} &= \frac{\tilde{\alpha}}{\bar{x}} = 0.000163.\end{aligned}$$

    -   the MMEs of $\mu$ and $\sigma$ for a
        $\mathcal{LN}(\mu, \sigma^2)$ distribution are $$\begin{aligned}
        \tilde{\sigma} &= \sqrt{ \ln \left(  \frac{s^2}{\bar{x}^2} + 1 \right)  }  = 1.393218, \\
        \tilde{\mu} &= \ln(\bar{x}) - \frac{\tilde{\sigma}^2 }{2} = 5.967012.\end{aligned}$$

    -   the MMEs of $\alpha$ and $\lambda$ for a
        $\text{Pa}(\alpha, \lambda)$ distribution are $$\begin{aligned}
        \tilde{\alpha} &= \displaystyle{ 2 \left(  \frac{s^2}{\bar{x}^2} \right) \frac{1}{(\frac{s^2}{\bar{x}^2} - 1)}   } = 2.402731,\\
        \tilde{\lambda} &= \bar{x} (\tilde{\alpha} - 1) = 1445.138.\end{aligned}$$

3.  The upper boundaries for the 10 groups or bins so that the expected
    number of claims in each bin is 20 under the fitted exponential
    distribution are determined by
    $$\Pr(X \le \text{upbd}_j) = \frac{j}{10}, \quad j = 1,2,3, \ldots, 9.$$
    With $\tilde{\lambda}$ from the MME for an $\text{Exp}(\lambda)$ from the previous,
    <!-- [\[example2\]](#example2){reference-type="ref" -->
    <!-- reference="example2"}, -->
    $$\Pr(X \le x)  = 1 - \exp(-\tilde{\lambda} x).$$ We obtain
    $$\text{upbd}_j = -\frac{1}{\tilde{\lambda}} \ln\left( 1 - \frac{j}{10}\right).$$
    The results are given in Table \@ref(tab:tableFitted). 

4.  The following table shows frequency distributions for observed and
    fitted claims sizes for exponential, gamma, and also lognormal and Pareto fits.


Table: (\#tab:tableFitted) Frequency distributions for observed and fitted claims sizes.
    
Range  | Observation |    Exp |  Gamma |  Lognormal |  Pareto
--: | --:|  --:|  --:| --: | --: |
(0,109]   |     60|          20|   109.4|          36|     31.9
(109,230] |     31|          20|    14.3|        34.4|     27.8
(230,367] |     25|          20|     9.7|          26|     24.2
(367,526] |     17|          20|     7.8|        20.5|     21.2
(526,714] |     14|          20|     6.8|        16.6|     18.6
(714,944] |     13|          20|     6.3|        13.9|     16.4
(944,1240]|      6|          20|     6.2|        11.9|     14.6
(1240,1658]|     7|          20|     6.5|        10.8|     13.2
(1658,2372]|     10|         20|     7.7|        10.4|     12.5
(2372,$\infty$)| 17|         20|    25.4|        19.5|     19.4



5.  Let $X$ be the claim size.

    -   The expected number of claims for the fitted exponential
        distribution in the range $(a,b]$ is
        $$200 \cdot \Pr( a < X \le b) = 200( e^{-\tilde{\lambda} a} - e^{-\tilde{\lambda} b} ).$$
        In our case, the expected frequencies under the fitted
        exponential distribution are given in the third column of Table \@ref(tab:tableFitted).
        <!-- [1](#tableFitted){reference-type="ref" reference="tableFitted"}. -->

    -   (Excel) The expected number of claims for the fitted gamma distribution
        in the range $(a,b]$ is
        $$200 \cdot\left(  \text{GAMMADIST}\left(b, \tilde{\alpha}, \frac{1}{\tilde{\lambda}}, \text{TRUE}\right)  - \text{GAMMADIST}\left(a, \tilde{\alpha}, \frac{1}{\tilde{\lambda}}, \text{TRUE}\right) \right).$$
        The expected frequencies under the fitted gamma distribution are
        given in the fourth column of Table \@ref(tab:tableFitted).

    -   (Excel) For the fitted lognormal, the expected number of claims in the
        range $(a,b]$ can be obtained from
        $$200 \cdot\left(  \text{NORMDIST} \left(\frac{LN(b) - 
                \tilde{\mu}}{\tilde{\sigma}}\right)  - \text{NORMDIST}\left(\frac{LN(a) - 
                \tilde{\mu}}{\tilde{\sigma}}\right) \right).$$

    -   For the fitted Pareto distribution, the expected number of
        claims in the range $(a,b]$ can be obtained from
        $$200 \left[    \left(\frac{\tilde{\lambda}}{\tilde{\lambda} + a} \right)^{\tilde{\alpha}}  - \left(\frac{\tilde{\lambda}}{\tilde{\lambda} + b} \right)^{\tilde{\alpha}}   \right].$$

6.  The histograms for the data set with fitted distributions
    are shown in Figures \@ref(fig:FittedExpGamma) and \@ref(fig:FittedLognormalPareto).
    <!-- [2](#FittedExpGamma){reference-type="ref" -->
    <!-- reference="FittedExpGamma"} and -->
    <!-- [3](#FittedLognormalPareto){reference-type="ref" -->
    <!-- reference="FittedLognormalPareto"}. -->

7.  Comments:

    1.  The high positive skewness of the sample reflects the fact that
        SD is large when compared to the mean. Consequently, the
        exponential distribution may not fit the data well.

    2.  Five claims (2.5%) are greater than 10,000, which is one of the
        main features of the loss distribution.

    3.  The fit is poor for the exponential distribution, as we see that
        the model under-fits the data for small claims up to 367 and
        over-fits for large claims between 944 to 2372. The gamma fit is
        again poor. We see that the model over-fits for small claims
        between 0-109 and under-fits for claims 230 and 944.

    4.  Which one of the lognormal and Pareto distributions provides a
        better fit to the observed claim data?

<!-- ![ **Histogram of claim sizes with fitted exponential and gamma -->
<!-- distributions.** ](Example3ClaimSizesExpoGamma.pdf){#FittedExpGamma -->
<!-- width="80%"} -->

<!-- ![ **Histogram of claim sizes with fitted lognormal and Pareto -->
<!-- distributions.** -->
<!-- ](Example3ClaimSizesParetoLognormal.pdf){#FittedLognormalPareto -->
<!-- width="80%"} -->


```{r, echo = FALSE}
dat <- c(31.08942156,915.0259362,32.02379562,885.875451,9314.100797,707.1738866,2114.362486,601.5827866,435.3788135,49.80179619,1802.38382,211.636439,153.5984719,60.05966992,48.17157697,948.7234626,132.4273109,151.710815,2967.961036,735.4149332,304.5807587,50.53311069,224.7665546,356.1809438,475.7668441,3044.149513,13.50664894,67.98274449,304.1439966,238.000105,399.0798263,149.28078,128.2848379,21.42027718,73.31426732,49.46882791,667.5327275,44.67360185,14894.07839,660.7614307,100.7628628,632.2812391,42.90884777,66.17693135,50.69454132,186.8781667,168.5408615,151.739978,2419.532354,434.4564032,69.03460232,162.6698593,226.6110395,33.61503495,233.027996,3244.945893,354.2534813,78.61052744,231.566614,283.0206491,457.6785426,134.2933058,61.34265063,38.25691912,1578.909048,441.6199826,761.4177777,275.6978588,521.0355916,2189.62783,117.2667885,240.3403153,672.4512938,753.8958019,84.98883079,3665.417976,60.35977515,4.01933309,14.95951366,196.7146403,153.7133165,98.31875053,104.4486324,358.9205878,2.254986331,2059.603959,37.63898608,56.48944035,1143.082948,410.7585516,12.65402986,19.83322514,1305.134797,2019.360735,1286.98479,8892.182131,5.814518749,296.1551295,86.75308161,486.7851405,7.491389798,180.3255281,1414.297748,524.4628208,1042.690334,1291.481474,115.4950998,360.6310737,3233.716838,149.5191039,8.845837473,83.87296324,42.99614517,623.9704853,45.74990074,144.2449793,368.575642,866.9272545,57.61592921,1812.231315,2229.998754,3448.332888,11313.34721,1492.498856,196.7262571,71.11817601,425.0614483,38.28653048,44.5090016,308.8781657,19080.51749,82.0761393,250.083523,79.07492042,183.8697686,33.83160391,22.78218299,698.9541648,32.75427003,457.0104919,112.1715567,396.7155234,195.0240781,1863.185385,181.6441121,59.08733655,96.39004191,824.730184,155.8215162,11.85510661,870.8679502,425.3133045,854.7296474,254.3108892,664.3320107,56.26120752,378.244016,2069.431495,3121.0901,844.4337594,746.3853675,1851.469633,431.7060232,333.325481,22.23970102,662.3816888,117.7590057,570.4412829,1127.035306,246.2686183,1467.568867,34.89187112,237.5761693,349.4503967,229.3362849,934.6926561,162.9254408,58.28497172,12803.04726,15.94904189,959.9543402,5853.878979,537.3974253,75.30572883,718.6597521,633.8214438,363.0338076,95.8476266,80.3178653,286.7121763,6367.453408,321.567711,21.51838022,599.4712958,246.7007073,13863.78181,214.7342597,234.3212952,958.916179,165.2521385)
dat <- data.frame(claims = dat)
```


```{r FittedExpGamma, tut = FALSE, fig.cap= 'Histogram of claim sizes with fitted exponential and gamma distributions.', echo = TRUE, warning= FALSE}

library(stats)
library(MASS)
library(ggplot2)

xbar <- mean(dat$claims)
s <- sd(dat$claims)

# MME of alpha and lambda for Gamma distribution
alpha_tilde <- (xbar/s)^2
lambda_tilde <- alpha_tilde/xbar

ggplot(dat) + geom_histogram(aes(x = claims, y = ..density..), bins = 90 , fill = "grey", color = "black") + 
  stat_function(fun=dexp, geom ="line", args = (rate = 1/mean(dat$claims)), aes(colour = "Exponential")) + 
  stat_function(fun=dgamma, geom ="line", args = list(shape = alpha_tilde ,rate = lambda_tilde), aes(colour = "Gamma")) + ylim(0, 0.0015) + scale_color_discrete(name="Fitted Distributions")
```

```{r FittedLognormalPareto, tut = FALSE, fig.cap= 'Histogram of claim sizes with fitted lognormal and pareto distributions.', echo = TRUE, warning= FALSE}
library(actuar)
# MME of mu and sigma for lognormal distribution

sigma_tilda <- sqrt(log(  var(dat$claims)/mean(dat$claims)^2 +1  ))  # gives \tilde\sigma
mu_tilda <- log(mean(dat$claims)) - sigma_tilda^2/2      # gives \tilde\mu

# MME of alpha and lambda for Pareto distribution
alpha_tilda <- 2*var(dat$claims)/mean(dat$claims)^2 * 1/(var(dat$claims)/mean(dat$claims)^2 - 1) #/tilde/alpha
lambda_tilda <- mean(dat$claims)*(alpha_tilda -1)


ggplot(dat) + geom_histogram(aes(x = claims, y = ..density..), bins = 90 , fill = "grey", color = "black") + 
  stat_function(fun=dlnorm, geom ="line", args = list(meanlog = mu_tilda, sdlog = sigma_tilda), aes(colour = "Lognormal")) + 
  stat_function(fun=dpareto, geom ="line", args = list(shape = alpha_tilda, scale = lambda_tilda), aes(colour = "Pareto")) +
  scale_color_discrete(name="Fitted Distributions")

```



Let us plot the histogram of claim sizes with fitted exponential and gamma distributions in this interaction area. Note that the data set is stored in the variable `dat`.

```{r ex="prob5", type="pre-exercise-code"}
library(stats)
library(MASS)
library(ggplot2)
dat <- c(31.08942156,915.0259362,32.02379562,885.875451,9314.100797,707.1738866,2114.362486,601.5827866,435.3788135,49.80179619,1802.38382,211.636439,153.5984719,60.05966992,48.17157697,948.7234626,132.4273109,151.710815,2967.961036,735.4149332,304.5807587,50.53311069,224.7665546,356.1809438,475.7668441,3044.149513,13.50664894,67.98274449,304.1439966,238.000105,399.0798263,149.28078,128.2848379,21.42027718,73.31426732,49.46882791,667.5327275,44.67360185,14894.07839,660.7614307,100.7628628,632.2812391,42.90884777,66.17693135,50.69454132,186.8781667,168.5408615,151.739978,2419.532354,434.4564032,69.03460232,162.6698593,226.6110395,33.61503495,233.027996,3244.945893,354.2534813,78.61052744,231.566614,283.0206491,457.6785426,134.2933058,61.34265063,38.25691912,1578.909048,441.6199826,761.4177777,275.6978588,521.0355916,2189.62783,117.2667885,240.3403153,672.4512938,753.8958019,84.98883079,3665.417976,60.35977515,4.01933309,14.95951366,196.7146403,153.7133165,98.31875053,104.4486324,358.9205878,2.254986331,2059.603959,37.63898608,56.48944035,1143.082948,410.7585516,12.65402986,19.83322514,1305.134797,2019.360735,1286.98479,8892.182131,5.814518749,296.1551295,86.75308161,486.7851405,7.491389798,180.3255281,1414.297748,524.4628208,1042.690334,1291.481474,115.4950998,360.6310737,3233.716838,149.5191039,8.845837473,83.87296324,42.99614517,623.9704853,45.74990074,144.2449793,368.575642,866.9272545,57.61592921,1812.231315,2229.998754,3448.332888,11313.34721,1492.498856,196.7262571,71.11817601,425.0614483,38.28653048,44.5090016,308.8781657,19080.51749,82.0761393,250.083523,79.07492042,183.8697686,33.83160391,22.78218299,698.9541648,32.75427003,457.0104919,112.1715567,396.7155234,195.0240781,1863.185385,181.6441121,59.08733655,96.39004191,824.730184,155.8215162,11.85510661,870.8679502,425.3133045,854.7296474,254.3108892,664.3320107,56.26120752,378.244016,2069.431495,3121.0901,844.4337594,746.3853675,1851.469633,431.7060232,333.325481,22.23970102,662.3816888,117.7590057,570.4412829,1127.035306,246.2686183,1467.568867,34.89187112,237.5761693,349.4503967,229.3362849,934.6926561,162.9254408,58.28497172,12803.04726,15.94904189,959.9543402,5853.878979,537.3974253,75.30572883,718.6597521,633.8214438,363.0338076,95.8476266,80.3178653,286.7121763,6367.453408,321.567711,21.51838022,599.4712958,246.7007073,13863.78181,214.7342597,234.3212952,958.916179,165.2521385)
dat <- data.frame(claims = dat)
```

```{r ex="prob5", type="sample-code"}
# Note that the object "dat" is data frame. To access the claims data we use dat$claims. # Fill your code below. 
class(dat)

```

```{r ex="prob5", type="solution"}
ggplot(dat) + 
  geom_histogram(aes(x = claims, y = ..density..), bins = 90 , fill = "grey", color = "black") + 
  stat_function(fun=dexp, geom ="line", args = (rate = 1/mean(dat$claims)), aes(colour = "Exp")) +
  scale_color_discrete(name="Fitted Distributions")
```





The following code can be used to obtain the expected number of claims for the fitted exponential distribution and perform goodness-of-fit test.




```{r ex="prob4", type="pre-exercise-code"}
library(stats)
library(MASS)
library(ggplot2)
dat <- c(31.08942156,915.0259362,32.02379562,885.875451,9314.100797,707.1738866,2114.362486,601.5827866,435.3788135,49.80179619,1802.38382,211.636439,153.5984719,60.05966992,48.17157697,948.7234626,132.4273109,151.710815,2967.961036,735.4149332,304.5807587,50.53311069,224.7665546,356.1809438,475.7668441,3044.149513,13.50664894,67.98274449,304.1439966,238.000105,399.0798263,149.28078,128.2848379,21.42027718,73.31426732,49.46882791,667.5327275,44.67360185,14894.07839,660.7614307,100.7628628,632.2812391,42.90884777,66.17693135,50.69454132,186.8781667,168.5408615,151.739978,2419.532354,434.4564032,69.03460232,162.6698593,226.6110395,33.61503495,233.027996,3244.945893,354.2534813,78.61052744,231.566614,283.0206491,457.6785426,134.2933058,61.34265063,38.25691912,1578.909048,441.6199826,761.4177777,275.6978588,521.0355916,2189.62783,117.2667885,240.3403153,672.4512938,753.8958019,84.98883079,3665.417976,60.35977515,4.01933309,14.95951366,196.7146403,153.7133165,98.31875053,104.4486324,358.9205878,2.254986331,2059.603959,37.63898608,56.48944035,1143.082948,410.7585516,12.65402986,19.83322514,1305.134797,2019.360735,1286.98479,8892.182131,5.814518749,296.1551295,86.75308161,486.7851405,7.491389798,180.3255281,1414.297748,524.4628208,1042.690334,1291.481474,115.4950998,360.6310737,3233.716838,149.5191039,8.845837473,83.87296324,42.99614517,623.9704853,45.74990074,144.2449793,368.575642,866.9272545,57.61592921,1812.231315,2229.998754,3448.332888,11313.34721,1492.498856,196.7262571,71.11817601,425.0614483,38.28653048,44.5090016,308.8781657,19080.51749,82.0761393,250.083523,79.07492042,183.8697686,33.83160391,22.78218299,698.9541648,32.75427003,457.0104919,112.1715567,396.7155234,195.0240781,1863.185385,181.6441121,59.08733655,96.39004191,824.730184,155.8215162,11.85510661,870.8679502,425.3133045,854.7296474,254.3108892,664.3320107,56.26120752,378.244016,2069.431495,3121.0901,844.4337594,746.3853675,1851.469633,431.7060232,333.325481,22.23970102,662.3816888,117.7590057,570.4412829,1127.035306,246.2686183,1467.568867,34.89187112,237.5761693,349.4503967,229.3362849,934.6926561,162.9254408,58.28497172,12803.04726,15.94904189,959.9543402,5853.878979,537.3974253,75.30572883,718.6597521,633.8214438,363.0338076,95.8476266,80.3178653,286.7121763,6367.453408,321.567711,21.51838022,599.4712958,246.7007073,13863.78181,214.7342597,234.3212952,958.916179,165.2521385)
dat <- data.frame(claims = dat)
```



```{r ex="prob4", type="sample-code"}
library(tidyr)
j = 0:9
upbd = qexp(j/10, 1/mean(dat$claims))  #1/mean(dat$claims) gives the parameter of exponential dist
observation <- cut(dat$claims, breaks = upbd,dig.lab=10)
table(observation)

# To fix the class interval (upbd when j = 9,infinity)
upbd[length(upbd)+1] = 1000000
upbd
observation <- cut(dat$claims, breaks = upbd)

tabSummary <- table(observation)

barplot(tabSummary,main="Claim sizes",las=2)

# Obtain E(Exp), the fitted claim sizes
eexp <- diff(200*pexp(upbd, rate = 1/mean(dat$claims)))

tabSummary <- data.frame(tabSummary) 
colnames(tabSummary) <- c("Range", "Observation")

tabSummary <- data.frame(tabSummary, Exponential = eexp)

#summary of exponential fit
chisq.test(x = tabSummary$Observation,
           p = eexp/sum(eexp))

# Manually compute chi-square statistics
sum((tabSummary$Observation - eexp)^2 /eexp)   
```



<!-- curve(dgamma(x ,shape = alpha_tilde ,rate = lambda_tilde), add = TRUE, col = "red", lwd = 2) -->




<!-- The following code creates plots of loss distributions in R using -->
<!-- ggplot2. See the following link for more information -->
<!-- <http://t-redactyl.io/blog/2016/03/creating-plots-in-r-using-ggplot2-part-9-function-plots.html> -->



<!-- ![ **The probability density functions (pdf) of exponential -->
<!-- distributions with various parameters $\lambda$.** -->
<!-- ](FigExp.eps){width="80%"} -->


<!-- The following code uses to obtain the results in -->
<!-- Example [Example 3](#exampleFittingClaimSizes){reference-type="ref" -->
<!-- reference="exampleFittingClaimSizes"}. -->

